Data Specialization for a Staged Lambda Calculus

### 2014-11-13 paper/splitting.tex
expansion of 5.4

\input{figures/splitting-overview}

@@

>>>>

@@

\subsection{Improved Definition}
\label{sec:resumers}

The definitions of masking and splitting can be augmented to better handle special cases,
particularly ones that depend on the type of the term we're splitting.
One such special case is the behavior of stage \bbone\ splitting at stage \bbone\ types that contain no $\fut$s or $\to$s.
We call such types {\em monostage} (judgmentally, $A~mono$), 
because a partial value $v$ with a monostage type can contain no $\next$s,
and so must comprise only stage \bbone\ information.

Intuitively, for $[\xi;v] : A$ and $A~mono$, 
we would expect that $\maskt{\xi;v}$ only reduce to something that carries no information.
This indeed happens: for example, $\maskt {\cdot;\mathtt{(inl~3,5)}}$ reduces to \texttt{((),())}.
But while \texttt{((),())} technically ``carries no information," it still has non-trivial structure.
This is not only unintuitive, but potentially expensive.
Instead, we desire a stronger property, that $\maskt{\xi;v}$ only reduce to exactly \texttt{()}. 
We can achieve this by adjusting just the tuple rule:
\[
\maskt{\cdot;(v_1,v_2)} = \left\{ \begin{array}{ll} 
\mathtt{()} &\text{if } (v_1,v_2):A~mono\\ 
(\maskt{\cdot;v_1},\maskt{\cdot;v_2}) &\text{otherwise} \end{array} \right.
\]
From here, the splitting rules can be made to comport with this new 
definition adding special-case tuple, projection, and application 
rules:
\begin{mathpar}
\infer
	{(e_1,e_2):A\times B\splitonesym [\ldots,(l_1,l_2).(r_1;r_2;())]}
	{e_1 : A \splitonesym [c_1,l_1.r_1] 
	& e_2 : B \splitonesym [c_2,l_2.r_2] 
	& A \times B~mono} 
\and
\infer
	{\pio~e:A\splitonesym [\ldots,l.(r;())]}
	{e : A\times B \splitonesym [c,l.r] 
	& A \times B~mono} 
\and
\infer
	{e_1~e_2: B\splitonesym [\ldots,(l_1,l_2,l_3).(r_1 (r_2,l_3); ())]}
	{e_1 : A \to B \splitonesym [c_1,l_1.r_1] 
	& e_2 : A \splitonesym [c_2,l_2.r_2] 
	& B~mono} 
\end{mathpar}
The old tuple, projection, and application rules 
must also be augmented to apply only when the $mono$ premises don't hold.
Also, all splitting rules need to be augmented to carry types, though
we omit the contexts here.
